Determinants of Wage Variation in Canada

### 2024-12-04 other/datasheet/datasheet.qmd
Finalize paper, update sketches and datasheet

1.  *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    -   The dataset was developed to investigate the relationship between education level, age, gender, and average hourly wage rates in Canada from 2000 to 2019. Its purpose is to assess whether higher educational attainment leads to increased earnings and how this relationship varies across demographic factors and has evolved over time.
2.  *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    -   Yuanyi (Leo) Liu, while studying at the University of Toronto.
3.  *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    -   No applicable.
4.  *Any other comments?*
    -   No additional comments are provided.

@@

1.  *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*
    -   The dataset represents aggregated hourly wage data categorized by education level, age group, and gender for individuals in Canada. Each instance corresponds to a unique combination of these factors, capturing variations in wages across demographic and educational categories from 2000 to 2019. The dataset does not include multiple types of instances beyond these categories.
2.  *How many instances are there in total (of each type, if appropriate)?*
    -   There are 1080 instances in the dataset, each representing a unique combination of year, education level, age group, gender, and average hourly wages.
3.  *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
    -   The dataset appears to be a complete set of instances from the available data during the specified timeframe. Representativeness is assumed but not specifically validated.

@@

8.  *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
    -   There are no recommended data splits mentioned, as the data appears to be used for comprehensive temporal analysis rather than predictive modeling that would require such splits.

@@

3.  *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
    -   As the dataset was derived from a comprehensive administrative record, the sampling strategy is not applicable. The data represents a census of the information available rather than a sample.

>>>>

1.  *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    -   The dataset was developed to investigate the relationship between education level, age, gender, and average hourly wage rates in Canada from 2000 to 2019. Its purpose is to assess whether higher educational attainment leads to increased earnings and how this relationship varies across demographic factors and has evolved over time.

2.  *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    -   Yuanyi (Leo) Liu, while studying at the University of Toronto.

3.  *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    -   No applicable.

@@

1.  *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*
    -   The dataset represents aggregated hourly wage data categorized by education level, age group, and gender for individuals in Canada. Each instance corresponds to a unique combination of these factors, capturing variations in wages across demographic and educational categories from 2000 to 2019. The dataset does not include multiple types of instances beyond these categories.

2.  *How many instances are there in total (of each type, if appropriate)?*
    -   There are 1080 instances in the dataset, each representing a unique combination of year, education level, age group, gender, and average hourly wages.

3.  *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
    -   The dataset appears to be a complete set of instances from the available data during the specified timeframe. Representativeness is assumed but not specifically validated.

@@

8.  *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
    -   There are no recommended data splits mentioned, as the data appears to be used for temporal analysis rather than predictive modeling that would require such splits.

@@

3.  *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
    -   As the dataset was derived from an administrative record, the sampling strategy is not applicable. The data represents a census of the information available rather than a sample.

### 2024-12-04 paper/paper.qmd
Finalize paper, update sketches and datasheet

@fig-education-level presents the distribution of average hourly wages in Canada from 2000 to 2019 across different education levels. Higher education levels, such as "University degree" and "Above bachelor's degree," are associated with higher median wages, while lower education levels, like "0–8 years," show significantly lower wages. The variability within each education level, shown by the spread of the boxes and whiskers, indicates differences in earning potential within these groups. For instance, individuals with an "Above bachelor's degree" earn higher wages on average and exhibit greater wage variability compared to those with "0–8 years" of education.

@@

@fig-gender compares the distribution of average hourly wages in Canada from 2000 to 2019 by gender. It displays the spread of wages for each gender, with males generally earning higher wages as shown by the broader and higher distribution. The jittered points represent individual wage observations, further emphasizing the overlap and differences in wage ranges. While both distributions have a similar shape, the median wage for males is noticeably higher than that for females, reflecting a persistent gender wage gap over the period analyzed.

@@

It also highlights differences in wages across age groups. Older individuals, particularly those aged 55 and over, earn higher wages compared to younger cohorts, reflecting the influence of experience and career progression. Additionally, the intersection of gender, age, and education reveals further disparities. For example, males with higher education levels in older age groups consistently exhibit the highest wages, while younger females with lower education levels earn the least.

@@

The analysis reveals a consistent wage disparity between males and females, with males earning much more across all education levels and age groups. This persistent wage gap suggests systemic inequities in how labor markets reward work. Factors contributing to this disparity may include occupational segregation, differences in work hours, career interruptions, and implicit biases in hiring and promotion processes.

>>>>

@fig-education-level presents the distribution of average hourly wages in Canada from 2000 to 2019 across different education levels. Higher education levels, such as "University degree" and "Above bachelor's degree," are associated with higher median wages, while lower education levels, like "0–8 years," show lower wages. The variability within each education level, shown by the spread of the boxes and whiskers, indicates differences in earning potential within these groups. For instance, individuals with an "Above bachelor's degree" earn higher wages on average and exhibit greater wage variability compared to those with "0–8 years" of education.

@@

@fig-gender compares the distribution of average hourly wages in Canada from 2000 to 2019 by gender. It displays the spread of wages for each gender, with males generally earning higher wages as shown by the broader and higher distribution. The jittered points represent individual wage observations, further emphasizing the overlap and differences in wage ranges. While both distributions have a similar shape, the median wage for males is higher than that for females, reflecting a persistent gender wage gap over the period analyzed.

@@

It also highlights differences in wages across age groups. Older individuals, particularly those aged 55 and over, earn higher wages compared to younger cohorts, reflecting the influence of experience and career progression. Additionally, the intersection of gender, age, and education shows further disparities. For example, males with higher education levels in older age groups consistently exhibit the highest wages, while younger females with lower education levels earn the least.

@@

The analysis presents a consistent wage disparity between males and females, with males earning much more across all education levels and age groups. This persistent wage gap suggests systemic inequities in how labor markets reward work. Factors contributing to this disparity may include occupational segregation, differences in work hours, career interruptions, and implicit biases in hiring and promotion processes.
